---
title: "Case Study 1"
author: "Richard Hart"
date: "February 16, 2019"
output: 
  html_document:
    keep_md: true
---

## Load Libraries
```{r, echo=T, message=F, warning=F}
# Load Libraries
library(cowplot)
library(plyr)
library(ggplot2)
```

# Read in and Setup Data
```{r, echo=TRUE}
# Create Data Frames from CSV files
beers <- read.csv(file="data/Beers.csv", stringsAsFactors=FALSE, header = TRUE, sep = ",")
breweries <- read.csv(file = "data/Breweries.csv", stringsAsFactors=FALSE, header = TRUE, sep = ",")

# Check the Class of each 
class(beers)
class(breweries)

# Check the structure of data frame
str(beers)
str(breweries)

# Change Brew_ID to Brewery_id so that both data frames have the same
# column name for Brewery ID
names(breweries)[1] <- "Brewery_id"

# Change Name in both data frame to a more descriptive column name
names(breweries)[2] <- "Brewery_name"
names(beers)[1] <- "Beer_name"
```

# Introduction

Breweries and Beers created, made, developed and loved in the United States. 

* include details about the beers, and breweries
* Mr. Customer the following is a detailed analysis answering the 7 questions of interest.
  
# Analysis

## Question of Interest 1 
### Number of Breweries in Each State
```{r, echo=T}
library(knitr)
kable(table(breweries$State), caption = "Table of Beers Per State", col.names = c("State", "Beer Count"), align = "l")
```

## Question of Interest 2 
### Merged Dataset, First 6 and Last 6 observations
```{r, echo=T}
beers_brews <- merge(beers, breweries, by="Brewery_id", all=TRUE)
head(beers_brews, 6)
tail(beers_brews, 6)
```

## Question of Interest 3 
### Report the number of NAs in each column
```{r, echo=T}
colSums(is.na(beers_brews))
```

## Question of Interest 4 
### Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
```{r, echo=T}
median_abv <- tapply(beers_brews$ABV, beers_brews$State, median, na.rm = T)
median_ibu <- tapply(beers_brews$IBU, beers_brews$State, median, na.rm = T)
```

#### Median Alcohol by Volume
```{r, echo=T}
median_abv
```

#### Median Internation Bitterness Units
```{r, echo=T}
median_ibu
```

#### create data frames from the above medians to be used in Graphs
```{r, echo=T, warning=F}
med_abv_df <- data.frame(template=names(median_abv), median=median_abv, stringsAsFactors = FALSE)
colnames(med_abv_df) <- (c("State", "ABV"))
med_ibu_df <- data.frame(template=names(median_ibu), median=median_ibu, stringsAsFactors = FALSE)
colnames(med_ibu_df) <- (c("State", "IBU"))
# Merge the two together
abv_ibu <- merge(med_abv_df, med_ibu_df, by="State")
```

#### Ditributions of ABV and IBU
```{r, echo=T, warning=F}
plot1 <- ggplot(abv_ibu, aes(x = State, y = ABV)) + 
  geom_bar(stat = "identity", fill="steelblue") +
  coord_flip() +
  xlab("State") + ylab("Median Alcohol By Volume") + 
  ggtitle("Distribution of Median ABV by State") +
  theme(axis.text.y = element_text(size = 5))
plot2 <- ggplot(abv_ibu, aes(x = State, y = IBU)) + 
  geom_bar(stat = "identity", fill="brown2") + 
  coord_flip() +
  xlab("State") + ylab("Median International Bitterness Units") + 
  ggtitle("Distribution of Median IBU by State") + 
  theme(axis.text.y = element_text(size = 5))
plot_grid(plot1, plot2)
```

```{r, echo=F}
detach("package:cowplot", unload = TRUE)
```
## Question of Interest 5 State with the most ABV and which with the most IBU

### State with the Highest ABV (ALL)
```{r, echo=T}
beers_brews[which.max(beers_brews$ABV), ]
```

##### Only State Abbreviation
```{r, echo=T}
beers_brews[which.max(beers_brews$ABV), ]$State
```

#### State with the Highest IBU (ALL)
```{r, echo=T}
beers_brews[which.max(beers_brews$IBU), ]
```

##### Only State Abbreviation
```{r, echo=T}
  beers_brews[which.max(beers_brews$IBU), ]$State
```

# Question of Interest 6 Summary Statistics of Alcohol by Volume
```{r, echo=T}
summary(na.omit(beers$ABV))
```

# Question 7 ScatterPlot of the Relationship Between IBU and ABV
```{r, echo=TRUE, warning=F}
pp<- ggplot(beers, aes(x=ABV, y=IBU)) +
  geom_point(aes(col=IBU, size=ABV)) +
  geom_smooth(method="loess", se=F) +
  xlim(c(0, 0.14)) + ylim(c(0, 140)) +
  labs(y = "International Bitterness Units",
       x = "Alcohol by Volume") +
  ggtitle("Relationship between International Bitterness Unit and Alcohol by Volume") +
  theme(plot.title = element_text(hjust=0.5)) +
  theme_gray()
plot(pp)
```
